---
title: "244 Lab Week 1"
author: "Kat Cotti"
date: "1/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set
library(tidyverse)
library(here) #Paths to read in data 
library(sf) #spatial data wrangling 
library(tmap) #interactive map 
library(dplyr)
library(tidyr)
```

### Read in SF trees data: 

```{r}
#Read in the data:
sf_trees <-
  read.csv(here("data", "sf_trees", "sf_trees.csv"))
```

### Basic Wrangling Reminders:

Refresh some skills for data wrangling and summary statistics using 'dplyr' & make a graph.

```{r}
top_five_status <- sf_trees %>%  
  count(legal_status) %>% #count function - replaces need for group_by by recognizing groups in the data, counts values (replaces summarize n), and puts in a table - count by legal status 
  drop_na(legal_status) %>% #removes any rows with missing or NA values for variable you specify - think about missing values before you remove them (206 final lecture)
  rename(tree_count = n) %>% # new name = old name  
  relocate(tree_count) %>% #moves tree count column to the front 
  slice_max(tree_count, n =5) #allows you to identify the rows with the highest value of the variable you specify and keep the top # of what you specify
```

Make a graph of those top 5 observations by legal status 

```{r}
ggplot(data = top_five_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count)) +
  geom_col() +
  labs(x = "Legal Status",
       y = "Tree Count") +
  coord_flip() + #flips the axis to fit in long title names 
  theme_minimal() 
# Default for ggplot is to plot in alphabetical order of category 
```

### A few more data wrangling refresher examples

Only want to keep observations (rows) for Blackwood Acacia Trees 

```{r}
blackwood_acacia <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>%  #look within a dataset and filter to obnly keep rows where within variable given the string is detected 
  select(legal_status, date, latitude, longitude) #picks columns you want to keep or exclude
 
## Make a faux map of the spatial orientation of this data using latitude and longitude:
ggplot(data = blackwood_acacia,
       aes(x = longitude,
            y = latitude)) +
  geom_point()
```

### Use 'tidyr' separate() and unite() functions 

Useful for combining OR separating columns ie. in this data set there are scientific names separated by :: to the common name 

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::") #separates the species column into scientific name and common name separated at :: 
```

Example: tidyr:: unite() - combine 2 columns into 1 column 
```{r}
sf_treess_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "_cool_") #new column name, colon indicates from this column to the next, what is going to be the separator 
```

### Make some actual maps of blackwood acacia trees in SF:

'st_as_sf()' to convert latitude and longitude to spatial coordinates. 

```{r}
#Convert latitude and longitude into spatial coordinates:
blackwood_acacia_sp <- blackwood_acacia %>% 
  drop_na(longitude,latitude) %>% #remove missing obs where lat and long are missing 
  st_as_sf(coords = c("longitude", "latitude"))

st_crs(blackwood_acacia_sp) = 4326 #set coordinate reference system 

ggplot(data = blackwood_acacia_sp) +
  geom_sf(color = "darkgreen")
```

Read in other data for the SF roads shapefile: 
```{r}
sf_map <- read_sf(here("data", "sf_map", "tl_2017_06075_roads.shp"))

# if there is an existing crs use st transform
st_transform(sf_map, 4326)

ggplot(data= sf_map) +
  geom_sf()
```

### Combine blackwood acacia tree observations & SF Roads Map:

```{r}
ggplot() +
  geom_sf(data = sf_map,
          size = 0.1, #changes size of the lines
          color = "darkgrey") + 
  geom_sf(data = blackwood_acacia_sp, #plots spatial points for blackwood acacia trees on top of the road map 
          color = "red",
          size = 0.5) +
  theme_void()
```

Make this an interactive map: still interactive in the html file. 

```{r}
tmap_mode("view") #default is plot which is a static map --> sets to interactive viewing

tm_shape(blackwood_acacia_sp) +
  tm_dots()
```

